---
title: "Association Rules Exercise"
author: "Sagar Chadha"
date: "8/17/2018"
output: 
  md_document:
    variant: markdown_github
  keep_md: yes
  fig_width: 12
  fig_height: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Grocery Store Purchase Analysis using Association Rules

Reading in the data, we find that the data has 9,835 transactions. Each row is one transaction and each comma separated item is one item from that transaction!

```{r, echo = FALSE, include = FALSE}
library(ggplot2)
library(dplyr)
library(ggthemes)
library(directlabels)
library(RColorBrewer)
library(ggalluvial)
library(tidyverse)
library(arules)
library(arulesViz)
library(tm) 
library(magrittr)
library(slam)
library(proxy)
library(xgboost)
```


```{r, echo = FALSE}
groceries <- read.table("Groceries.txt", header = FALSE, sep = ";", stringsAsFactors = FALSE)

```

 Let's look at one row from our data -

```{r}
groceries[1,]
```

Currently our data is just comma separated items with each row of our data being one transaction. For our analysis we need to convert this to a list with each element of the list being a vector with the items from a transaction.

```{r}
item_list <- list()

for (i in 1:nrow(groceries)){
  item_list[[i]] <- unlist(strsplit(groceries[i,], ","))
}

item_list <- apply(groceries, MARGIN = 1, FUN = function(x){unlist(strsplit(x, ","))})

item_vector <- unlist(item_list)
```

### What is the number of different items sold here?

This would give us a sense of what customers are buying from this store and lead into our analysis of what items are most frequently bought together.

```{r, echo = FALSE}
length(unique(item_vector))

```

### What sells most commonly?

```{r, echo = FALSE}
count_item <- as.data.frame(table(item_vector))

ggplot(count_item %>% 
         arrange(-Freq) %>% 
         head(20), 
       aes(x = reorder(item_vector, -Freq), y = Freq)) + 
  geom_bar(stat = "identity") + 
  coord_flip() +
  scale_x_discrete("Times sold") +
  scale_y_continuous("Item Name")
```

The most commonly sold item is **Whole Milk** followed by **Vegetables** and **Rolls/Buns**! Out of the 169 unique items sold, we see that only 15 odd items are sold in considerable amounts. 

### There is a long tail of low selling items! What are these?

```{r}
count_item %>%
  arrange(Freq) %>%
  head(10)

```

### What items are associated with other items on the grocery list? 

Association here means items that are most commonly sold together in a transaction! We will use a low _minsup_ (minimum support) threshold here since we want item associations that would not be obvious to the grocer at first glance- Items like milk, eggs, vegetables are the most commonly bought items and most people know about their associations.

```{r}
# consider only unique items in each basket
item_list <- lapply(item_list, unique)

# convert to the transactions type
item_transactions <- as(item_list, "transactions")

# apply the apriori
groceryrules <- apriori(item_transactions,
                        parameter=list(support = .001,
                                       confidence = .5))
head(groceryrules, n = 10, by ="lift")

plot(head(groceryrules, n = 10, by ="lift"), method = "graph", 
     main = "Top 10 Association Rules")
```

### Conclusion - 
A study of the associations shows us the following - <br>

1. People purchase soda, popcorn and other salty snacks together.
2. Cheese, ham, white bread and eggs usually sell together.
3. Sugar, baking powder and flour sell together, these are usually baking items.
4. Cheese, curd, whipped cream and yogurt sell together!